---
title: "An Analysis of Chicago's Divvy Transport"
author: "Imogen Meers, Sarah Deussing, Sarah Cernugel"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

#### Questions
We want to look at the trends of Divvy bike and scooter data in Chicago, Illinois and how Divvy use changes with events that occur throughout the year. Doing so will allow us to recommend business changes to improve Divvy's revenue in Chicago. 

#### Data Cleaning & Preprocessing
Our data comes from many different sources. We will use Divvy data from 2020-2024, as well as professional sports schedules from the same years. These sports schedules will come from R packages, and our important columns will be game date and game start time. To merge all sports schedule data sets, we will have to standardize the format of these dates and times.

We will then merge this data with the Divvy data to look at changes to Divvy use on these pre-identified games and times. In terms of creating new values, we can look at the change/difference in Divvy use from days without sporting events to days with games. We will also look for patterns in vehicle type and membership, which will be useful to provide recommendations for where to adding new docks or make more bikes/scooters available depending on the event and if there are patterns between casual users vs members.

## Initial Data Analysis
Data Cleaning, Exploratory Visualizations

#### Divvy Data
```{r}
divvy <- read.csv("divvy data.csv")
head(divvy)
```

```{r}
library(dplyr)
library(readr)
library(purrr)
library(stringr)
library(tibble)
library(magick)
library(aws.s3)
library(tidyr)
library(ggplot2)
library(ggrepel)
```

```{r Reading from S3 bucket, eval = FALSE}
bucket_exists("s3://divvy-tripdata/") #check bucket exists

data <- get_bucket_df(
  bucket = "s3://divvy-tripdata/", 
  region = "us-east-1", 
  max = 20000
) %>% 
  as_tibble() #find contents of bucket

keys <- data[1:54,]$Key #get zip files that we want

save_object(object = keys[[1]], bucket = "divvy-tripdata",region = "us-east-1", file = temp) 
s3read_using(FUN = save_object, object = keys[[1]], bucket = "divvy-tripdata", region = "us-east-1", file = temp)

download_and_unzip <- function(key) { 
  temp <- tempfile() 
  save_object(object = key, bucket = "divvy-tripdata",region = "us-east-1", file = temp) 
  unzipped_files <- unzip(temp, exdir = tempdir()) 
  print(paste("File ", key, " unzipped"))
  return(read.csv(unzipped_files[1]))
} # Function to unzip each file into a temp file 

# Apply function to each key in your data frame 
result_list <- lapply(keys, download_and_unzip)

# Combine into a single data frame 
result_df <- do.call(rbind, result_list)


# save_object(
#   object = "202004-divvy-tripdata.zip",
#   bucket = "s3://divvy-tripdata/", 
#   region = "us-east-1",
#   file = "202004-divvy-tripdata.zip"
# )

#Should save as a RData file to prevent having to run this every time
```

```{r Reading from S3 bucket}
load("InitRData.RDataTmp")
```

```{r}
summary(result_df)
head(result_df)
```

```{r}
library(dplyr)
head(result_df)
unique(result_df$rideable_type)
summary(as.factor(result_df$rideable_type))

scooter <- result_df %>% filter(rideable_type == "electric_scooter")
scooter_sample <- sample_n(scooter, 5000)

summary(scooter)
classic_bike <- result_df %>% filter(rideable_type == "classic_bike")
classic_sample <- sample_n(classic_bike, 5000)

docked_bike <- result_df %>% filter(rideable_type == "docked_bike")
docked_sample <- sample_n(docked_bike, 5000)

electric_bike <- result_df %>% filter(rideable_type == "electric_bike")
electric_sample <- sample_n(electric_bike, 5000)

member_sample <- result_df %>% filter(member_casual == "member") %>% sample_n(5000)
casual_sample <- result_df %>% filter(member_casual == "casual") %>% sample_n(5000)

unique(result_df$member_casual)
head(result_df)
```

```{r}
library(ggmap)
register_stadiamaps("9e07144d-cdab-48f9-a35d-298e1bbece2e")

# Define the bounding box for Chicago
bbox <- c(left = -87.9400, bottom = 41.6445, right = -87.5240, top = 42.0230)

# Get the map
chi_map <- get_stadiamap(bbox = bbox, zoom = 12, maptype = "stamen_toner")


# Plot the map
scooter_map <- ggmap(chi_map) +
  geom_point(data = scooter_sample, aes(x = start_lng, y = start_lat, alpha = 0.05), color = "red") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
  ) + 
  labs(
    title = "Scooter Map"
  )+
  guides(alpha = "none")

scooter_map
```

```{r}
# Plot the map
classic_map <- ggmap(chi_map) +
  geom_point(data = classic_sample, aes(x = start_lng, y = start_lat, alpha = 0.05), color = "blue") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
  ) + 
  labs(
    title = "Classic Bike Map"
  )+
  guides(alpha = "none")

docked_map <- ggmap(chi_map) +
  geom_point(data = docked_sample, aes(x = start_lng, y = start_lat, alpha = 0.05), color = "green") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
  ) + 
  labs(
    title = "Docked Bike Map"
  )+
  guides(alpha = "none")


electric_map <- ggmap(chi_map) +
  geom_point(data = electric_sample, aes(x = start_lng, y = start_lat, alpha = 0.05), color = "orange") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
  ) + 
  labs(
    title = "Electric Bike Map"
  )+
  guides(alpha = "none")

classic_map
electric_map
docked_map


library(gridExtra)
grid.arrange(scooter_map, classic_map, electric_map, docked_map, nrow = 2, ncol = 2)
```
```{r}
# Plot the map
member_map <- ggmap(chi_map) +
  geom_point(data = member_sample, aes(x = start_lng, y = start_lat, alpha = 0.05), color = "blue") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
  ) + 
  labs(
    title = "Member Map"
  )+
  guides(alpha = "none")

casual_map <- ggmap(chi_map) +
  geom_point(data = casual_sample, aes(x = start_lng, y = start_lat, alpha = 0.05), color = "green") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
  ) + 
  labs(
    title = "Casual Map"
  )+
  guides(alpha = "none")

library(gridExtra)
grid.arrange(casual_map, member_map, ncol=2)
```

## NBA Schedule
```{r}
#install.packages("hoopR")
library(hoopR)
nba_19 <- nba_schedule(league_id = "00", season = 2019)
nba_20 <- nba_schedule(league_id = "00", season = 2020)
nba_21 <- nba_schedule(league_id = "00", season = 2021)
nba_22 <- nba_schedule(league_id = "00", season = 2022)
nba_23 <- nba_schedule(league_id = "00", season = 2023)
nba_24 <- nba_schedule(league_id = "00", season = 2024)

#nba <- full_join(nba_20, full_join(nba_21, full_join(nba_22, full_join(nba_23, nba_24))))
nba <- rbind(nba_19,nba_20, nba_21, nba_22, nba_23, nba_24)

nba <- nba %>% filter(arena_name == "United Center" & game_status == 3)
nba %>% arrange(desc(game_date))
summary(nba)
nba_dates <- unique(nba$game_date)
```

## MLB Schedule
```{r}
#install.packages("baseballr")
library(baseballr)

mlb_19 <- mlb_schedule(season = 2019, level_ids = "1")
mlb_20 <- mlb_schedule(season = 2020, level_ids = "1")
mlb_21 <- mlb_schedule(season = 2021, level_ids = "1")
mlb_22 <- mlb_schedule(season = 2022, level_ids = "1")
mlb_23 <- mlb_schedule(season = 2023, level_ids = "1")
mlb_24 <- mlb_schedule(season = 2024, level_ids = "1")

mlb <- rbind(mlb_19, mlb_20, mlb_21, mlb_22, mlb_23, mlb_24)
#mlb <- full_join(mlb_20, full_join(mlb_21, full_join(mlb_22, full_join(mlb_23, mlb_24))))

mlb_wrigley <- mlb %>% filter(venue_name == "Wrigley Field") %>% mutate(date = as.Date(date))
wrigley_dates <- unique(mlb_wrigley$date)

mlb_grfield <- mlb %>% filter(venue_name == "Guaranteed Rate Field") %>% mutate(date = as.Date(date))
grfield_dates <- unique(mlb_grfield$date)

mlb <- mlb %>% filter(venue_name == "Wrigley Field" | venue_name == "Guaranteed Rate Field") %>% mutate(date = as.Date(date))
summary(mlb)
mlb_dates <-  unique(mlb$date)

summary(as.Date(mlb_24$date))
```



## NFL Schedule
```{r}
#install.packages("nflfastR")
library(nflfastR)
nfl <- nflreadr::load_schedules()
nfl$gameday <- as.Date(nfl$gameday)
nfl <- nfl %>%
  filter(gameday >= as.Date('2020-01-01') & gameday <= as.Date('2024-12-31') & stadium == "Soldier Field")
summary(nfl)
nfl_dates <-  unique(nfl$gameday) 
```
## Cleaning NFL Schedule
```{r}
pbp <- load_pbp(2019:2024)
pbp <- pbp %>% data.frame() %>% filter(home_team == "CHI") 
pbp <- pbp[!is.na(pbp$time_of_day),]
pbp_start <- pbp[!duplicated(pbp$game_id),]
pbp_temp <- pbp[nrow(pbp):1,]
pbp_end <- pbp_temp[!duplicated(pbp_temp$game_id),]


pbp_end <- pbp_end %>%  select(game_date, time_of_day) %>% rename(end_time =time_of_day)
pbp_start <- pbp_start %>%  select(game_date,time_of_day)%>% rename(start_time=time_of_day)
soldier_schedule <- merge(x = pbp_start, y = pbp_end, by = c("game_date"))

soldier_schedule$start_time <- as.POSIXct(soldier_schedule$start_time,format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC")
soldier_schedule$start_time <- time_at_tz(soldier_schedule$start_time, tz = "America/Chicago")
 
 
soldier_schedule$end_time <- as.POSIXct(soldier_schedule$end_time,format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC")
soldier_schedule$end_time <- time_at_tz(soldier_schedule$end_time, tz = "America/Chicago")
soldier_schedule$game_date <- as.Date(soldier_schedule$game_date)

inner_join(soldier_schedule, nfl, by = c("game_date"="gameday"))
```

```{r}
head(soldier_schedule)
head
```

# NHL Schedule
```{r}
#install.packages('fastRhockey')
library(fastRhockey)
nhl19 <- load_nhl_schedule(2019)
nhl20 <- load_nhl_schedule(2020)
nhl21 <- load_nhl_schedule(2021)
nhl22 <- load_nhl_schedule(2022)
nhl23 <- load_nhl_schedule(2023)
nhl24 <- load_nhl_schedule(2024)

nhl <- rbind(nhl19, nhl20, nhl21, nhl22, nhl23, nhl24, fill=TRUE)
nhl <- nhl %>% filter(venue_name == "United Center") %>% mutate(game_date = as.Date(game_date))
summary(nhl)
nhl_dates <-  unique(nhl$game_date)
```

Combine into all Dates
```{r}
dates <- seq.Date(from = as.Date("2020-01-01"), to = as.Date("2024-12-31"), by = "day")
dates <- data.frame(date = dates)

mlb_dates <- as.Date(mlb_dates)
nba_dates <- as.Date(nba_dates)
nfl_dates <- as.Date(nfl_dates)
nhl_dates <- as.Date(nhl_dates)
dates <- dates %>%
  mutate(mlb = if_else(date %in% mlb_dates, 1, 0),
         nba = if_else(date %in% nba_dates, 1, 0),
         nfl = if_else(date %in% nfl_dates, 1, 0),
         nhl = if_else(date %in% nhl_dates, 1, 0))
```

Event "Calendar" By Sport
```{r}
library(ggplot2)
library(tidyr)

dates_long <- dates %>%
  pivot_longer(cols = c(mlb, nhl, nba, nfl), 
               names_to = "sport", 
               values_to = "game", 
               names_prefix = "_events") %>%
  filter(game == 1)

dates_long <- dates_long %>%
  mutate(year = as.integer(format(date, "%Y")))

ggplot(dates_long, aes(x = date, y = sport, color = sport)) +
  geom_point(size = 3) +
  scale_y_discrete(limits = c("mlb", "nhl", "nba", "nfl")) +
  labs(x = "Date", y = "Sport", title = "Game Occurences by Sport") +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_wrap(~ year, scales = "free_x", ncol = 1)
```

Total Events Per Day
```{r}
events_per_day <- dates_long %>%
  group_by(date) %>%
  summarise(events = n()) %>%
  mutate(year = as.integer(format(date, "%Y")))

ggplot(events_per_day, aes(x = date, y = events)) +
  geom_line(color = "blue", size = 1) +
  labs(x = "Date", y = "Number of Events", title = "Sports Events Per Day") +
  theme_minimal() +
  facet_wrap(~ year, scales = "free_x", ncol = 1) + 
  theme(legend.position = "none")
```

```{r Filter to game days}
filtered_df <- result_df %>% mutate(start_date = as.Date(started_at)) %>% filter(start_date %in% dates_long$date)
```

```{r}
load("FilteredAndTeams.RData")
summary(filtered_df)
filtered_df <- filtered_df %>% na.omit()
```
## Venues
```{r Venue Radius}
library(stringr)
library("timechange")
# Define the custom function
round_and_adjust <- function(number, adjust_by = 0.005) {
  # Round the number to three decimal places
  rounded_number <- round(number, 3)
  
  # Adjust the rounded number by the specified amount
  adjusted_number <- rounded_number + adjust_by
  
  return(adjusted_number)
}

venue_df <- read.csv("chicago_venues.csv")
venue_df$max_lat <- round_and_adjust(venue_df$Latitude, 0.005)
venue_df$min_lat <- round_and_adjust(venue_df$Latitude, -0.005)
venue_df$max_lng <- round_and_adjust(venue_df$Longitude, 0.005)
venue_df$min_lng <- round_and_adjust(venue_df$Longitude, -0.005)

venue_df$League <- str_extract(venue_df$Team, "(?<=\\().*(?=\\))")
venue_df$Team <- str_extract(venue_df$Team, "^[^(]+")
venue_df <- venue_df %>% select(1:3, League, everything())
venue_df
write.csv(venue_df, "chicago_venues.csv")
```


## Cleaning Baseball Schedule
```{r}
game_times_wrigley <- read_csv("game_times.csv")

clean_baseball <- function(game_times_wrigley, mlb_wrigley){
  game_times_wrigley$Date <- as.Date(strptime(game_times_wrigley$Date, format = "%A, %b %d %Y"))



joined_wrigley <- inner_join(mlb_wrigley, game_times_wrigley, by = c("date" = "Date"))


joined_wrigley$game_date <- as.POSIXct(joined_wrigley$game_date,format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC")
joined_wrigley$local_start_time <- time_at_tz(joined_wrigley$game_date, tz = "America/Chicago")

joined_wrigley$local_end_time <- joined_wrigley$local_start_time + joined_wrigley$Time


wrigley_schedule <- joined_wrigley %>% select(date, Attendance, Time, local_start_time, local_end_time, season, game_date, game_number, day_night, series_game_number, series_description, teams_away_team_name,venue_name, description, teams_away_score, teams_away_is_winner,teams_home_score,teams_home_is_winner )

return(wrigley_schedule)
}
game_times_gr <- read_csv("game_times_whitesox.csv")

wrigley_schedule <- clean_baseball(game_times_wrigley, mlb_wrigley)
grfield_schedule <- clean_baseball(game_times_gr, mlb_grfield)
```


## Wrigley Field Graph
```{r}
wrigley_df <- result_df %>% filter((start_lat <= venue_df$max_lat[1] & start_lat >= venue_df$min_lat[1] & start_lng <= venue_df$max_lng[1] & start_lng >= venue_df$min_lng[1] )|(end_lat <= venue_df$max_lat[1] & end_lat >= venue_df$min_lat[1] & end_lng <= venue_df$max_lng[1] & end_lng >= venue_df$min_lng[1]) ) %>% mutate(start_date = as.Date(started_at)) %>% mutate(gameday = ifelse(start_date %in% wrigley_dates, 1, 0))

plot_data<- wrigley_df %>% select(start_date, gameday) %>% group_by(start_date, gameday) %>% mutate(count = n()) %>% unique()


outliers <- plot_data %>% filter(gameday == 0) %>% distinct() %>%arrange(desc(count)) %>% head(8)

outliers$desc <- c("Dead & Co Concert", "", "Zac Brown Band Concert", "Dead & Co Concert", "P!NK Concert", "Chicago Pride", "", "Belmont Sheffield Music Fest")

wrigley_graph <- ggplot() + 
  geom_point(data = plot_data, aes(x = start_date, y = count, color = as.factor(gameday))) +  geom_text_repel(data = outliers, aes(x = start_date, y = count, label = outliers$desc), color = "red", size = 3) + labs(title = "Count of Divvy Bikes in Wrigley Field Radius")

wrigley_graph
```
```{r}
head(wrigley_df)
summary(wrigley_df)
library(lubridate)
wrigley_df <-  wrigley_df %>% mutate(ride_time = difftime(ended_at, started_at,units = "mins"))


##Look at distribution of ride times
hist_plot <- wrigley_df[wrigley_df$gameday==1 & wrigley_df$member_casual == "casual",]
time_plot <- c(as.numeric(hist_plot$ride_time %>% na.omit()))
time_plot <- time_plot[time_plot> 0]
is_outlier <- abs(time_plot - mean(time_plot)) > 2 * sd(time_plot)
time_plot <- time_plot[!is_outlier]
summary(hist_plot$ride_time)

cas <- hist(time_plot, breaks = 30, main = "Histogram of Ride Times of Divvy Users", xlab = "ride_time", ylab = "Frequency", col = "darkgreen", border = "black") ##look to se what the distribution of divvy users ride time is, looks exponentail


hist_plot <- wrigley_df[wrigley_df$gameday==1 & wrigley_df$member_casual == "member",]
time_plot <- c(as.numeric(hist_plot$ride_time %>% na.omit()))
time_plot <- time_plot[time_plot> 0]
is_outlier <- abs(time_plot - mean(time_plot)) > 2 * sd(time_plot)
time_plot <- time_plot[!is_outlier]
summary(hist_plot$ride_time)

mem <- hist(time_plot, breaks = 30, main = "Histogram of Ride Times of Divvy Users", xlab = "ride_time", ylab = "Frequency", col = "darkgreen", border = "black") ##look to se what the distribution of divvy users ride time is, looks exponentail


library(gridExtra)
grid.arrange(plot(cas), plot(mem), nrow=1)

rate <- 1/mean(time_plot)


summary(time_plot)

wrigley_schedule$season <- as.numeric(wrigley_schedule$season)


attendance_impute <- mean(wrigley_schedule[season == 2021,]$Attendance, na.rm=TRUE) #only the 2020 had missing data so impute with the 2021 season


wrigley_schedule <- wrigley_schedule %>% mutate(Attendance = ifelse(is.na(Attendance), round(attendance_impute), Attendance)) #impute NA attendance

master_wrigley <- inner_join(wrigley_schedule, wrigley_df, by= c("date" ="start_date")) #join the schedule and rides


####################### whole day

member_casual_perc <- master_wrigley %>% group_by(date, season, member_casual) %>% summarise(count = n()) %>% mutate(perc = count/sum(count))

rider_count_per_game <- master_wrigley %>% group_by(date, season) %>% summarise(count = n()) #summarise num of riders per game


rider_count_per_game <- merge(x = (master_wrigley %>% select(date, Attendance)), y = rider_count_per_game, by = "date") %>% distinct() %>% mutate(prob_rider = count/Attendance)

expected_games_season <- rider_count_per_game %>% group_by(season) %>% summarise(expected_rider_games = sum(prob_rider))

expected_games_season
#######################

################################ 3 hours + or - game
master_wrigley <- master_wrigley %>% mutate(game_rider = ifelse(started_at >= local_start_time - hours(3) & started_at <= local_end_time + hours(3) | ended_at >= local_start_time - hours(3) & ended_at <= local_end_time + hours(3), 1, 0 )) #mark every classed as a game rider (ends or starts in radius on game day )

member_casual_perc <- master_wrigley %>% group_by(date, season, member_casual) %>% summarise(count = n()) %>% mutate(perc = count/sum(count))

rider_count_per_game <- master_wrigley %>% group_by(date, season) %>% summarise(sum(game_rider)) #summarise num of riders per game


rider_count_per_game <- merge(x = (master_wrigley %>% select(date, Attendance)), y = rider_count_per_game, by = "date") %>% distinct() %>% mutate(prob_rider = `sum(game_rider)`/Attendance)


expected_games_season <- rider_count_per_game %>% group_by(season) %>% summarise(expected_rider_games = sum(prob_rider))

expected_games_season
##############################

################## Simulation of different
summary(rider_count_per_game)



#times, calculate normal distribution mean and sd
#attendace and numebr of riders assuming 75% ride in and out calculate probs and expected number of games attended divvy bike

#sample calc profit with pricing (old and new)
#show price 
```

```{r}

plot_data <- wrigley_df %>% group_by(start_date) %>% mutate(total_count = n())%>% group_by(start_date, member_casual) %>% mutate(count_member = n()) %>% arrange(start_date) %>% select(start_date, gameday, count_member, member_casual, total_count) %>% distinct() %>% mutate(percent = count_member/total_count)

gameday <- plot_data %>% filter(gameday == 0)
no_gameday <- plot_data %>% filter(gameday == 1)
plot2023 <- plot_data %>% filter(start_date > as.Date("2023-01-01"))

wrigley_graph1 <- ggplot(plot2023, aes(x = start_date, y = count_member, fill = member_casual)) + geom_bar(stat = "identity") + labs(title = "Stacked Bar Chart")


head(wrigley_df)
```

## Guarantee Rate Field Graph
```{r}
grfield_df <- result_df %>% filter((start_lat <= 41.835 & start_lat >= 41.825 & start_lng <= -87.63 & start_lng >= -87.64 )|(end_lat <= 41.835 & end_lat >= 41.825 & end_lng <= -87.63 & end_lng >= -87.64) ) %>% mutate(start_date = as.Date(started_at)) %>% mutate(gameday = ifelse(start_date %in% grfield_dates, 1, 0))

plot_data<- grfield_df %>% select(start_date, gameday) %>% group_by(start_date, gameday) %>% mutate(count = n()) %>% unique()


outliers <- plot_data %>% filter(gameday == 0) %>% distinct() %>%arrange(desc(count)) %>% head(8)

outliers$desc <- rep(NA, 8)

grfield_graph <- ggplot() + 
  geom_point(data = plot_data, aes(x = start_date, y = count, color = as.factor(gameday))) +  geom_text_repel(data = outliers, aes(x = start_date, y = count, label = outliers$desc), color = "red", size = 3) + labs(title = "Count of Divvy Bikes in Guarantee Rate Field Radius")

grfield_graph
```
## United Center Graph
```{r United}
united_dates <- unique(c(nba_dates, nhl_dates))

united_df <- result_df %>% filter((start_lat <= 41.885 & start_lat >= 41.875 & start_lng <= -87.67 & start_lng >= -87.68 )|(end_lat <= 41.885 & end_lat >= 41.875 & end_lng <= -87.63 & end_lng >= -87.64) ) %>% mutate(start_date = as.Date(started_at)) %>% mutate(gameday = ifelse(start_date %in% nba_dates, "nba", ifelse(start_date %in% nhl_dates, "nhl", "none")))

# united_df <- result_df %>% filter((start_lat <= 41.885 & start_lat >= 41.875 & start_lng <= -87.67 & start_lng >= -87.68 )|(end_lat <= 41.885 & end_lat >= 41.875 & end_lng <= -87.63 & end_lng >= -87.64) ) %>% mutate(start_date = as.Date(started_at)) %>% mutate(gameday = ifelse(start_date %in% united_dates, 1, 0))

plot_data<- united_df %>% select(start_date, gameday) %>% group_by(start_date, gameday) %>% mutate(count = n()) %>% unique()

# outliers <- plot_data %>% filter(gameday == 0) %>% distinct() %>%arrange(desc(count)) %>% head(8)

outliers <- plot_data %>% filter(gameday == "none") %>% distinct() %>%arrange(desc(count)) %>% head(8)

outliers$desc <- rep(NA, 8)

united_graph <- ggplot() + 
  geom_point(data = plot_data, aes(x = start_date, y = count, color = as.factor(gameday))) +  geom_text_repel(data = outliers, aes(x = start_date, y = count, label = outliers$desc), color = "red", size = 3) + labs(title = "Count of Divvy Bikes in United Center Field Radius")

# united_graph1 <- ggplot() + 
#   geom_point(data = plot_data, aes(x = start_date, y = count, color = as.factor(gameday))) +  geom_text_repel(data = outliers, aes(x = start_date, y = count, label = outliers$desc), color = "red", size = 3) + labs(title = "Count of Divvy Bikes in United Center Field Radius")
united_graph
```

```{r}

soldier_df <- result_df %>% filter((start_lat <= 41.87 & start_lat >= 41.86 & start_lng <= -87.61 & start_lng >= -87.62 )|(end_lat <= 41.87 & end_lat >= 41.86 & end_lng <= -87.61 & end_lng >= -87.62) ) %>% mutate(start_date = as.Date(started_at)) %>% mutate(gameday = ifelse(start_date %in% nfl_dates, 1, 0))

# united_df <- result_df %>% filter((start_lat <= 41.885 & start_lat >= 41.875 & start_lng <= -87.67 & start_lng >= -87.68 )|(end_lat <= 41.885 & end_lat >= 41.875 & end_lng <= -87.63 & end_lng >= -87.64) ) %>% mutate(start_date = as.Date(started_at)) %>% mutate(gameday = ifelse(start_date %in% united_dates, 1, 0))

plot_data<- soldier_df %>% select(start_date, gameday) %>% group_by(start_date, gameday) %>% mutate(count = n()) %>% unique()

outliers <- plot_data %>% filter(gameday == 0) %>% distinct() %>%arrange(desc(count)) %>% head(3)


outliers$desc <- c("Coldplay Concert", "AC Milan vs. Real Madrid CF", "Chicago Fire FC vs.\n Inter Miami CF")

soldier_graph <- ggplot() + 
  geom_point(data = plot_data, aes(x = start_date, y = count, color = as.factor(gameday))) +  geom_text_repel(data = outliers, aes(x = start_date, y = count, label = outliers$desc), color = "red",segment.color = 'red' ,size = 3) + labs(title = "Count of Divvy Bikes in Soldier Field Radius")
soldier_graph

```

